---
title: "Pokemon Lab"
author: "Anish Yakkala"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
paperwidth <- 8.3
textwidth <- 11.7
```


```{r include=FALSE}
library(tidyverse)
library(Hmisc)
library(plotly)
library(forcats)
```

I scraped Smogon rankings using a script I wrote in Python (https://github.com/ayakkala1/Stat-331/blob/master/scrape_smogon.py): 

```{r}
pokes <-read.csv("https://www.dropbox.com/s/i0lwxgv86eaoq4o/pokemon.csv?dl=1")
smogon <- read.csv("/Users/ramanyakkala/Stat-331/smogon.csv")
```

# Pokemon Strength Analysis

Since I don't have domain knowledge on what makes a Pokemon strong I decided to use what competitive pokemon players use.

You can watch the most recent finals here (https://www.youtube.com/watch?v=ObGnEFujuBE).


This analysis will judge "strong pokemon" by using Smogon rankings. This is the go to place for competitive Pokemon players.

"Smogon University, commonly shortened to Smogon, is a website whose content encompasses competitive Pokémon battling. It offers guides on battling strategies for people of different Pokémon knowledge backgrounds. The website was founded by 'chaos' (one of the developers of NetBattle) and is a considerably well-known website, visited by competitive Pokémon battling enthusiasts."
  - https://bulbapedia.bulbagarden.net/wiki/Smogon

Let's clean the smogon rankings I scraped

```{r}
rankings <- smogon %>%
  filter(formats != "") %>%
  rename("Name" = "name")

```

Let's do Difference of Means tests of various fighting metrics by Uber vs Not Uber. (https://www.smogon.com/dex/sm/formats/uber/)

```{r warning=FALSE}
pokes %>%
  left_join(select(rankings,Name,formats, evos), by = "Name") %>%
  mutate(formats = replace_na(formats,"Untiered")) %>%
  mutate(Uber = ifelse(formats == "Uber",1,0)) %>%
  select_if(is.numeric) %>%
  select(-Number) %>%
  gather(key = variable, value = value, -Uber) %>%
  nest(-variable) %>%
  group_by(variable) %>%
  mutate(p_value = unlist(map(data, ~t.test(value ~ Uber,na.rm = TRUE,.)$p.value))) %>%
  arrange(p_value) %>%
  mutate(significant = ifelse(p_value < 0.05,"yes","no"))
```

Everything is significant, but also we have only two pokemon that are Uber ranking in our Gen 1-6 dataset. This is because most of the Uber's
are mega pokemon, various versions of legendary pokemon (ex: Arceus & Deoxys), or Gen 7.

Let's look at the OverUsed tier which is the next ranking. This is also the more common format for competitive pokemon.

```{r warning=FALSE}
pokes %>%
  left_join(select(rankings,Name,formats, evos), by = "Name") %>%
  mutate(formats = replace_na(formats,"Untiered")) %>%
  filter(formats != "Uber" ) %>%
  mutate(OU = ifelse(formats == "OU",1,0)) %>%
  select_if(is.numeric) %>%
  select(-Number) %>%
  gather(key = variable, value = value, -OU) %>%
  nest(-variable) %>%
  group_by(variable) %>%
  mutate(p_value = unlist(map(data, ~t.test(value ~ OU,na.rm = TRUE,alternative="less",.)$p.value)),
         t_value = unlist(map(data, ~t.test(value ~ OU,na.rm = TRUE,alternative="less",.)$statistic))) %>%
  arrange(p_value) %>%
  mutate(significant = ifelse(p_value < 0.05,"yes","no"))

```

Seems like all the one sided differences are siginficant, except for Weight & Generation & Base Catch Rate.

Those that are significant give good insights on what metrics matter for which Pokemon are in the OverUsed rankings.

Let's Graph the Differences.
```{r warning=FALSE}
pokes %>%
  left_join(select(rankings,Name,formats, evos), by = "Name") %>%
  mutate(formats = replace_na(formats,"Untiered")) %>%
  filter(formats != "Uber" ) %>%
  mutate(OU = ifelse(formats == "OU",1,0)) %>%
  select_if(is.numeric) %>%
  select(-Number) %>%
  group_by(OU) %>%
  summarize_all(funs(mean)) %>%
  gather(key,value,-OU) %>%
  ggplot(aes(x = as.character(OU),y = value)) + geom_col() + facet_wrap(~key, scales = "free_y") + 
    ggtitle("Comparisons of fighting metrics of OU vs not OU") + xlab("Non OU vs OU") + ylab("values") 
```

You can also see that the distrubution of types within formats is not homogeneous

Read up on formats here (https://www.smogon.com/tiers/)

```{r, fig.width=20,fig.height=10}
rankings %>%
  mutate(types = paste(Type1,Type2,sep =":")) %>%
  separate_rows(types, sep = ":", convert = FALSE) %>%
  filter(types != "") %>%
  filter(formats != "Limbo") %>%
  select(formats,types) %>%
  mutate(types = fct_infreq(types)) %>%
  ggplot(aes(x = types, fill = types)) + geom_bar(aes(y = ..prop..,group = 1))  + facet_wrap(~formats) + theme(axis.text.x = element_text(face="bold",size=14, angle=90)) + ggtitle("Type Distribution by Format")

```

------------------------------------------------------------------------------------------------------------------------------------------

# Fun Stuff

Some fun stuff I was doing with probability of catching a pokemon,before I realized that this lab had to do with finding stronger pokemon.

Was playing around with Plotly

This all started out by seeing the negative correlation between base catch rate and total stats. This makes sense as the stronger the pokemo is the harder it is to catch it!

```{r}
cor(x = pokes$Catch_Rate,y = pokes$Total)
```

Obviosly legendary pokemon are going to be really hard to catch, so let's get rid of them.

Let's look at base catch rate by types
```{r, fig.width=6,fig.height=4}
legendary <- c("Articuno","Zapdos","Moltres","Mewtwo","Mew","Raikou",
               "Entei","Suicune","Lugia","Ho-Oh","Latias","Latios","Kyogre",
               "Groudon","Rayquaza","Jirachi","Deoxys","Regirock","Regice",
               "Registeel","Uxie","Mesprit","Azelf","Dialga","Palkia","Heatran",
               "Regigigas","Giratina","Cresselia","Phione","Manaphy","Darkrai",
               "Shaymin","Arceus","Victini","Cobalion", "Virizion", "Tornadus",
               "Thundurus","Reshiram","Zekrom","Landorus","Kyurem","Keldeo",
               "Genesec","Xerneas","Yveltal","Zygarde","Diancie","Hoopa","Volcanion")

pokes %>%
  filter(!Name %in% legendary) %>%
  mutate(types = paste(Type_1,Type_2,sep=":")) %>%
  separate_rows(types, sep = ":", convert = FALSE) %>%
  filter(types != "") %>%
  mutate(Type = fct_reorder(types,Catch_Rate,.fun=mean)) %>%
  group_by(Type) %>%
  summarise(mean_catch = mean(Catch_Rate)) %>%
  mutate(Type = fct_reorder(Type,mean_catch,fun=n,desc=TRUE)) %>%
  ggplot(aes(x=Type, y=mean_catch, fill = Type)) + geom_col() + ylab("Average Base Catch Number") +ggtitle("Average Base Catch Number by Type")


```

I used Gen III-IV catch-rate formula and treated the pokeball as a regular pokeball, as well as that the pokemon is not under any status effect. (https://bulbapedia.bulbagarden.net/wiki/Catch_rate). I didn't use the further generations since it involved type of grass.

```{r, fig.width=6,fig.height=4}
# Modified catch rate
capture_k_trials <- function(shake_prob, k){
  
  prob_pass_shake <- (shake_prob-1)/65535
  
  prob_shake_fails <- 1-dbinom(4, size = 4, prob = prob_pass_shake)
  
  prob_one_succ <- 1-(prob_shake_fails)^k
    
  return(prob_one_succ)
}

modified_catch <- function(perc,HP,Catch_Rate){
  mod <- ((((3 * HP) - (2 * (HP*(perc/100))))*Catch_Rate)/(3*HP))
  return(mod)
}

current = pokes$HP %/% 2
pokes %>%
  mutate(mod_catch = ((3 * HP - 2 * current)*Catch_Rate)/(3*HP)) %>%
  mutate(shake_prob = 1048560 %/% floor(sqrt(floor(sqrt(16711680%/%mod_catch))))) %>%
  mutate(prob_succ = map_dbl(shake_prob,~capture_k_trials(.x,5))) %>%
    filter(!Name %in% legendary) %>%
    mutate(types = paste(Type_1,Type_2,sep=":")) %>%
    separate_rows(types, sep = ":", convert = FALSE) %>%
      filter(types != "") %>%
      mutate(Type = fct_reorder(types,prob_succ,.fun=mean)) %>%
        group_by(Type) %>%
        summarise(mean_catch = mean(prob_succ),mean_catch2 = mean(Catch_Rate)) %>%
        mutate(Type = fct_reorder(Type,mean_catch,fun=n,desc=TRUE)) %>%
            ggplot(aes(x=Type, y=mean_catch, fill = Type)) + geom_col() + ylab("Avg. Prob. of 1 success in 5 throws") + ggtitle("Average Probability of Successful Catch in 5 throws by Pokemon Type")

  
```

Let's try to see how probability of catching a Gible changes as its %health missing changes.

```{r}
pokes %>%
    filter(Name == "Gible") %>%
    select(HP,Catch_Rate) %>%
    pmap_dfr(
    ., 
    ~as.list(set_names(
        modified_catch(...), 
        paste0(1:100)
    )), 
    perc = 1:100
    ) %>%
    gather(key = "percent", value="mod_catch") %>%
    mutate(shake_prob = 1048560 %/% floor(sqrt(floor
                                               (sqrt(16711680%/%mod_catch))))) %>%
    mutate(prob_succ = map_dbl(shake_prob,~capture_k_trials(.x,1))) %>%
    mutate(percent = as.integer(percent)) %>%
    ggplot(aes(x=percent,y=prob_succ))+geom_point()+ylim(0,0.2) + ggtitle("Probability of Success, the higher the percent the Pokemon (Gible)") + xlab("Percent of Health of Pokemon") + ylab("Probability of Success")

```

Let's add on that by seeing how it changes as you keep throwing more pokeballs (Change Gible's % Health missing using the slider)
```{r}

Bagon <- list()
for(i in 1:100) {
  Bagon[[i]] <- list(visible = FALSE,
                     name = paste0('Percent of health missing: ',i),
                     x=1:20,
                     y= as.list(pokes %>%
                       filter(Name == "Gible") %>%
                       mutate(mod_catch = modified_catch(i, HP, Catch_Rate)) %>%
                       mutate(shake_prob = 1048560 %/% floor(sqrt(floor
                                               (sqrt(16711680%/%mod_catch)))))  %>%
                       select(shake_prob) %>%
                       pmap_dfr(
                       ., 
                       ~as.list(set_names(
                           capture_k_trials(...), 
                           paste0(1:20)
                       )), 
                       k = 1:20
                       ) %>%
                       gather(key = "tries",value="chance") %>%
                       select(chance)))
}

Bagon[100][[1]]$visible = TRUE

steps <- list()
p <- plot_ly()
for (i in 100:1) {
  p <- add_lines(p,x=Bagon[i][[1]]$x,  y=Bagon[i][[1]]$y$chance, visible = Bagon[i][[1]]$visible, 
                 name = Bagon[i][[1]]$name, type = 'scatter', mode = 'lines', hoverinfo = 'name', 
                 line=list(color='00CED1'), showlegend = FALSE) %>%
       layout(title = "Probability of Catching Gible as Health Changes",xaxis = list(title = 'K tries'),
                yaxis = list(title = 'Percent of Success'),
                legend = list(x = 0.80, y = 0.90))

  step <- list(args = list('visible', rep(FALSE, length(Bagon))),
               method = 'restyle')
  step$args[[2]][i] = TRUE  
  steps[[i]] = step 
}

p <- p %>%
  layout( 
          sliders = list(list(active = 100,
                             currentvalue = list(prefix = "Percentage Health Missing: "),
                             steps = steps))
              )

p
```

Was trying to work with showing probability of catching a pokemon using dropdowns. Could not get it to work with all pokemon, however notice how all the starter's and their evos have the same probability. This is because they are uncatchable pokemon!

```{r}
format_poke_graph <- function(i){
                return(list(method = "restyle",
                args = list("transforms[0].value", unique(test$Name)[i]),
                label = unique(test$Name)[i]))
}

test <- pokes %>%
         mutate(mod_catch = modified_catch(50, HP, Catch_Rate)) %>%
         mutate(shake_prob = 1048560 %/% floor(sqrt(floor
                                   (sqrt(16711680%/%mod_catch)))))  %>%
         select(shake_prob) %>%
         pmap_dfr(
                  ., 
                 ~as.list(set_names(
                  capture_k_trials(...), 
                  paste0(1:20)
                  )), 
                   k = 1:20
                  ) %>%
        add_column(Name = pokes$Name) %>%
        gather("tries","percent_succ",-Name) %>%
        mutate(tries = as.integer(tries))



steps <- list()
p <- plot_ly(data = test, x = ~tries, y = ~percent_succ)

p <- test %>%
  plot_ly(
    type = 'scatter',
    x = ~tries,
    y = ~percent_succ,
    text = ~Name,
    hoverinfo = 'text',
    mode = 'line',
    transforms = list(
      list(
        type = 'filter',
        target = ~Name,
        operation = '=',
        value = unique(test$Name)[1]
      )
    )) %>% layout(
        title = "Pick a pokemon",
        updatemenus = list(
          list(
            type = 'dropdown',
            active = 0,
            buttons = map(1:length(pokes),format_poke_graph)
    )))
  
p
```




